Effects of epidemiological structure on the transient evolution of HIV virulence

### 2016-10-19 ms/plos_latex.tex
added 1 todo in ms and worked on other todo's

\begin{equation}
\begin{split}
D_A(\alpha) &= \frac{\tsub{D}{max} D_{50} ^{D_k}}{V_\alpha ^{D_k} + D_{50}^{D_k}}, \\
\beta_A(\alpha) &= \frac{\tsub{\beta}{max} V_\alpha ^ {\beta_k}}%
{V_\alpha^{\beta_k} + \beta_{50} ^{\beta_k}},
\end{split}
\label{eq:hillfuns}
\end{equation}
where $V_{\alpha} = 10^\alpha$. 
The uncoupled and extra-couple transmission rates are scaled by multiplying the within-couple transmission rate $\beta$ by the contact ratios $c_u/c_w$ and $c_e/c_w$ (see Appendix). Simplifying the model of HIV pathogenesis from three stages to a single stage could affect our conclusions about the evolution of virulence (e.g. Kretzschmar and Dietz \cite{kretzschmar_effect_1998} show that pair formation dynamics and multiple stages of infectivity have interactive effects on $\rzero$). However, our simplified model produces results that are qualitatively similar to those of Shirreff \etal's \cite{shirreff_transmission_2011} model; when our model is calibrated to have a similar initial epidemic growth rate $r$, the peak \Lspvl\ occurs at the same time ($\approx$ 200 years) but slightly higher (4.7 vs. 4.3); the relative height of the peak is \todo{compute difference and fill this in}
(\figurename~\ref{fig:panel3}).

@@

\todo{add something about choice of heterogeneity}

@@

Across the entire range of parameters covered by the LHS analysis, all of the classes of models we considered produce qualitatively similar virulence trajectories (\figurename~\ref{fig:virtraj}). Although the speed of virulence evolution varies, leading to wide variation in the minimum progression time (means ranging from approximately \todo{fill in}), virulence peaks in all models between 200 and 300 years.

@@

These differences are practically as well as 
scientifically important. The random-mixing and pairform+epc models
predict minimum times to progression (at the virulence peak) of
5.7 (95\% CI 6.1-6.3) and 6.0 (5.0-7.7) years respectively, while
the implicit model gives progression times about twice as long:
12.5 (9.5-15.6) years. The corresponding differences in 
within-couple transmission
probability are even more extreme, about a fourfold difference:
0.249 (0.24-0.26) and 0.252 (0.19-0.28) per year for the 
random and pairform+epc models vs. 0.059 (0.02-0.13) per year
for the implicit model (Figs~???? \todo{fill in!} show univariate distributions
on the epidemiological scales
of progression time and transmission probability,
for all summary measures and all models).

>>>>

\begin{equation}
\begin{split}
D_A(\alpha) &= \frac{\tsub{D}{max} D_{50} ^{D_k}}{V_\alpha ^{D_k} + D_{50}^{D_k}}, \\
\beta_A(\alpha) &= \frac{\tsub{\beta}{max} V_\alpha ^ {\beta_k}}%
{V_\alpha^{\beta_k} + \beta_{50} ^{\beta_k}},
\end{split}
\label{eq:hillfuns}
\end{equation}
where $V_{\alpha} = 10^\alpha$. 
The uncoupled and extra-couple transmission rates are scaled by multiplying the within-couple transmission rate $\beta$ by the contact ratios $c_u/c_w$ and $c_e/c_w$ (see Appendix). Simplifying the model of HIV pathogenesis from three stages to a single stage could affect our conclusions about the evolution of virulence (e.g. Kretzschmar and Dietz \cite{kretzschmar_effect_1998} show that pair formation dynamics and multiple stages of infectivity have interactive effects on $\rzero$). However, our simplified model produces results that are qualitatively similar to those of Shirreff \etal's \cite{shirreff_transmission_2011} model; when our model is calibrated to have a similar initial epidemic growth rate $r$, the peak \Lspvl\ occurs at the same time ($\approx$ 200 years) but slightly higher (4.7 vs. 4.3); the relative height of the peak is 1.07 (\figurename~\ref{fig:panel3}).

@@

Two parameters, mean and the squared coefficient of variation (CV) of number of non-cohabiting partners, are sampled for heterogeneity in sexual activity. We decided to use squared CV instead of variance as variance measures is often unreliable \todo{CITE? You mentioned that variance is unreliable and we should use CV but we didn't talk about it very much...}. To allow for a wide range of uncertainty, range for the mean number of non-cohabiting partners was taken from unmarried men as that was the group with the largest variability \cite{omori2015dynamics}. However, we decided to use 0.01 - 100 as the sampling range for squared CV despite the wider CV range given by Omori \etal \cite{omori2015dynamics} (0 - 20.324). Although married female have largest variability for CV, the range of the mean is narrow and small, which might explain for such large variability. Assessing ranges of CV of other groups, we concluded that 0.01 - 100 is both reasonable and wide enough to cover uncertainties in parameters. For simplicity, we assume that the number of non-cohabiting partner follows a gamma distribution.

@@

Across the entire range of parameters covered by the LHS analysis, all of the classes of models we considered produce qualitatively similar virulence trajectories (\figurename~\ref{fig:virtraj}). Although the speed of virulence evolution varies, leading to wide variation in the minimum expected progression time (means ranging from approximately 6 to 12 years), virulence peaks in all models between 200 and 300 years.

@@

These differences are practically as well as 
scientifically important. The random-mixing and pairform+epc models
predict minimum times to progression (at the virulence peak) of
5.7 (95\% CI 6.1-6.3) and 6.0 (5.0-7.7) years respectively, while
the implicit model gives progression times about twice as long:
12.5 (9.5-15.6) years. The corresponding differences in 
within-couple transmission
probability are even more extreme, about a fourfold difference:
0.249 (0.24-0.26) and 0.252 (0.19-0.28) per year for the 
random and pairform+epc models vs. 0.059 (0.02-0.13) per year
for the implicit model (\figurename~\ref{fig:unidist} show univariate distributions
on the epidemiological scales
of progression time and transmission probability,
for all summary measures and all models).
